Understanding \& Developing Artificial Neural Networks

### 2017-03-03 research paper/paper.tex
Updated research paper. Double check reference and direct quotations.

%%use packages%%
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[export]{adjustbox}
\usepackage{lipsum}
\usepackage{titlesec}
\usepackage{lmodern}
\usepackage{titling}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{multicol}
\usepackage[a4paper, total={6in, 9in}]{geometry} %margin
\usepackage{fixltx2e} % provides \textsubscript
\usepackage{listings}
\usepackage{color}
\usepackage{caption}
\usepackage[]{microtype}
\usepackage{graphicx}
\usepackage{wrapfig}
\graphicspath{ {images/}}

@@

%define subtitle
\newcommand{\subtitle}[1]{%
  \posttitle{%
    \par\end{center}
    \begin{center}\large#1\end{center}
    \vskip0.5em}%
}

@@

There are many methods and algorithms for machine learning, from Support Vector
Machine\footnote{https://en.wikipedia.org/wiki/Support\_vector\_machine}
to Artificial Neural Networks\footnote{https://en.wikipedia.org/wiki/Artificial\_neural\_network}.
One of the common purpose for learning algorithms is to classify complex data. It can be used in analyzing human genetics and genomics. \footnote{Machine learning applications in genetics and genomics, Libbrecht, Maxwell W. http://dx.doi.org/10.1038/nrg3920}

@@

Neural networks, with their remarkable ability to derive meaning from complicated or imprecise data, can be used to extract patterns and detect trends that are too complex to be noticed by either humans or other computer techniques. A trained neural network can be thought of as an "expert" in the category of information it has been given to analyze. \footnote{This section is reference from NEURAL NETWORKS 
	by Christos Stergiou and Dimitrios Siganos  https://www.doc.ic.ac.uk/~nd/surprise\_96/journal/vol4/cs11/report.html\#Why\%20use\%20neural\%20networks}

@@

The neural network will begin be taking in some inputs and making some
predictions based on the inputs and the weights between nodes. We can
think of it like a one by two matrix. There will be a weight connecting
every input layer node to every output layer node, therefore, there are
six weights between the input layer and the second layer. The matrix
calculation should yield us some result

\[\begin{bmatrix}
a
\end{bmatrix}\ \lbrack w \rbrack = \begin{bmatrix}
z
\end{bmatrix}\]

If we give the matrices some names, call inputs \emph{x}, weights
\emph{w\textsubscript{(l)}} and output \emph{z\textsubscript{(l)}.} In
our case, \emph{l} indicate the layer, for example, the first layer
weights are \emph{w\textsubscript{(1)}}. \emph{z\textsubscript{(l)}}
represent the output matrix with layer \emph{l}, in this case, the
output z is the hidden layer output.

@@

Afterwards, we have to apply an activation function\footnote{https://en.wikipedia.org/wiki/Activation\_function}.
The activation function that I used in this research and this paper will
be the sigmoid function. This is a complete cycle, there is one more
layer to go in order to yield a result. Note, in a multilayer neural
network, this process will be repeated until we yield some output. In
our case, we only have to do this twice. The activation function is
shown as below.

\[\sigma\left( z \right) = \ \frac{1}{1 + e^{- z}}\]
\begin{equation}
	{\text{\ a}}_{\left( 2 \right)} = f(z_{\left( 2 \right)})\
\end{equation}

By applying the same process as before, this time, the inputs will be
\emph{a\textsubscript{(2)} then, z\textsubscript{(3)}} will be our final
result. This process is known as forward feed. It's relatively
straightforward.

\[z_{\left( 3 \right)} = \ \text{\ a}_{\left( 2 \right)}w_{(2)}\]

After we yield some result, we can more on and look at how far off we
are from the expected result. We need some methods to quantify this and
so adjustments to the network to minimize error.

\begin{equation}
	\sigma'\left( z \right) = \ \sigma\left( z \right)(1 - \sigma(z))\
\end{equation}

@@

C is the cost, which equals to the sum of all the differences between
calculated result and actual result squared and times one half. We can
take advantage of the equation derived above and substitute for some of
the variable.

\begin{equation}
	\ C = \ \sum_{J}^{}{\frac{1}{2}\left( y - f\left( f\left( xw_{\left( 1 \right)} \right)w_{\left( 2 \right)} \right) \right)^{2}}
\end{equation}

Here we have it, a way to quantify the cost of the neural network. This
function will be referred to as the cost function. Now, we have to solve
the problem, how do we minimize C, will brute force work? It turns out,
no, because in a three-node neural network we have to compute more than
a million possible weights, which will be gruesome.

@@

Gradient descent~is a~first-order~iterative~optimization~algorithm. To
find a~local minimum~of a function using gradient descent, one takes
steps proportional to the gradient (or of the
approximate gradient) of the function at the current point.

Gradient descent is also known as~steepest descent, or the~method of
steepest descent. The process can be seen as a ball rolling down a
hill\footnote{\url{https://iamtrask.github.io/2015/07/27/python-network-part2/}
  by Andrew}
and trying to find the lowest point. Note that actual physics doesn't
apply here and we will define our own movement of the ball.

@@

\emph{Phase 1: Propagation: Each propagation
involves the following steps:}
\footnote{The steps are referenced from
Wikipedia who referenced:
\href{http://numericinsight.com/uploads/A_Gentle_Introduction_to_Backpropagation.pdf}{A
Gentle Introduction to Backpropagation - An intuitive tutorial by
Shashi Sathyanarayana}~The article contains pseudocode ("Training
Wheels for Training Neural Networks") for implementing the algorithm.}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Forward propagation of a training pattern's input through the neural
  network in order to generate the network's output value(s).
\item
  Backward propagation of the propagation's output activations through
  the neural network using the training pattern target in order to
  generate the deltas (the difference between the targeted and actual
  output values) of all output and hidden neurons.
\end{enumerate}

\emph{
Phase 2: Weight update: For each weight, the following
steps must be
followed:}

@@

Backpropagation is based around four fundamental equations. Together,
those equations give us a way of computing both the error
\(\delta^{L}\)and the gradient cost of the function. In fact, the
backpropagation equations are so rich that understanding them well
requires considerable times and patience. \footnote{The equations were
  created by Michael A. Neilson "Neural Networks and Deep Learning",
  Determination Press, 2015~}

@@

The first term on the right, \(\frac{\partial J}{\partial a_{J}^{L}}\)
measures how fast the cost is changing as a function of
j\textsuperscript{th} output activation. Everything in (BP1) is easily
calculated. We computed \(z_{J}^{L}\) while computing the behavior of
the network. Depending on the cost function, in our case, the quadratic
cost function is relatively easy to compute.

Equation (BP1) is a perfectly good expression, however, it's not matrix
based, form that backpropagation desires. The fully matrix form becomes.

\begin{equation}
	\delta^{L} = (a^{l} - y)(\sigma'(z^{l})\
\end{equation}

@@

where \(\left( w^{l + 1} \right)^{T}\) is the transpose of the weight
matrix \(\left( w^{l + 1} \right)\) for the
\emph{(l+1)\textsuperscript{th}} layer. Suppose we know the error
\(\delta^{l + 1}\) at the \emph{(l+1)\textsuperscript{th }}layer. When
the transpose weight matrix is applied \(\left( w^{l + 1} \right)^{T}\),
we can think of this as moving the error backward through the network,
giving us some sort of measure of the error at the output of the
l\textsuperscript{th} layer. Finally, we take the Hadamard
product\(\ \bigodot\sigma'(z^{l})\).

By combining (BP1) with (BP2), the error \(\delta^{l}\) for any
layer in the network can by computed. We start by using~(BP1)~to
compute~\(\delta^{l}\), then apply Equation~(BP2)~to
compute~\(\delta^{L - 1}\), then Equation~(BP2)~again to
compute~\(\delta^{L - 1}\), and so on, all the way back through the
network.

@@

\subsubsection{An equation for the rate of change of the cost with respect to any weights in the network}

\begin{equation} \tag{BP4}
	\frac{\partial C}{\partial w^l_{jk}} = a^{l-1}_{k} \delta^l_{j}\footnote{The equation was
		referenced from Michael A. Neilson "Neural Networks and Deep Learning",
		Determination Press, 2015~}
\end{equation}

This equation can help us to compute the partial derivative \(\frac{\partial C}{\partial w^l_{jk}}\) in terms of the quantity \(\delta^l\) and \(a^{l-1}\). It can be rewritten in a less index-heavy notation:

\begin{equation}
	\frac{\partial C}{\partial w} = a_{in}\delta_{out}\
\end{equation}

\subsection{Backpropagation with simple neural network example}

>>>>

%%use packages%%
\usepackage[utf8]{inputenc}
\usepackage [english]{babel}
\usepackage [autostyle, english = american]{csquotes}
\usepackage[export]{adjustbox}
\usepackage{lipsum}
\usepackage{titlesec}
\usepackage{lmodern}
\usepackage{titling}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{multicol}
\usepackage[a4paper, total={6in, 9in}]{geometry} %margin
\usepackage{fixltx2e} % provides \textsubscript
\usepackage{listings}
\usepackage{color}
\usepackage{caption}
\usepackage[]{microtype}
\usepackage{graphicx}
\usepackage{wrapfig}
\graphicspath{ {images/}}

@@

\MakeOuterQuote{"}
%define subtitle
\newcommand{\subtitle}[1]{%
  \posttitle{%
    \par\end{center}
    \begin{center}\large#1\end{center}
    \vskip0.5em}%
}

@@

There are many methods and algorithms for machine learning, from Support Vector
Machine\footnote{https://en.wikipedia.org/wiki/Support\_vector\_machine}
to Artificial Neural Networks\footnote{https://en.wikipedia.org/wiki/Artificial\_neural\_network}.
One of the common purpose for learning algorithms is to classify complex data. For example, It can be used in analyzing human genetics and genomics. \footnote{Machine learning applications in genetics and genomics, Libbrecht, Maxwell W. http://dx.doi.org/10.1038/nrg3920}

@@

According to Christos Stergiou and Dimitrios Siganos' Neural Networks\footnote{This section is reference from NEURAL NETWORKS 
	by Christos Stergiou and Dimitrios Siganos  https://www.doc.ic.ac.uk/~nd/surprise\_96/journal/vol4/cs11/report.html\#Why\%20use\%20neural\%20networks} "Neural networks, with their remarkable ability to derive meaning from complicated or imprecise data, can be used to extract patterns and detect trends that are too complex to be noticed by either humans or other computer techniques. A trained neural network can be thought of as an "expert" in the category of information it has been given to analyze."

@@

The neural network will begin be taking in some inputs and making some
predictions based on the inputs and the weights between nodes. We can treat the inputs and weights as matrices. This process can be seen as consecutive matrices multiplications. The matrix
calculation should yield us some result

\[\begin{bmatrix}
input
\end{bmatrix}\ \lbrack weights \rbrack = \begin{bmatrix}
output
\end{bmatrix}\]

If we give the matrices some names, call inputs \emph{x}, weights
\emph{w\textsubscript{(l)}} and product of the matrix multiplication \emph{z\textsubscript{(l)}.} In our case, \emph{l} indicate the layer, for example, the first layer
weights are \emph{w\textsubscript{(1)}}. \emph{z\textsubscript{(1)}}
represent the output matrix with layer \emph{1}.

@@

Afterwards, we have to apply an activation function\footnote{https://en.wikipedia.org/wiki/Activation\_function}.
The activation function that I used in this research and this paper will
be the sigmoid function.

\[\sigma\left( z \right) = \ \frac{1}{1 + e^{- z}}\]
\begin{equation}
{\text{\ a}}_{\left( l \right)} = \sigma(z_{\left( l \right)})\
\end{equation}

Multiplying \(a_{previous}\) with connection weights, then applying activation function is one complete cycle. In a multilayer neural
network, this process will be repeated until we reach the output layer. In
a simple three layer neural network, this process only have to execute twice.

@@

C is the cost, which equals to the sum of all the differences between
calculated result and actual result squared and times one half.

In a simple three layer neural network, we can
take advantage of the equation derived above and substitute for some of
the variable.
\[\hat{y} = f\left( f\left( xw_{\left( 1 \right)} \right)w_{\left( 2 \right)} \right)\]
\begin{equation}
	\ C = \ \sum_{J}^{}{\frac{1}{2}\left( y - f\left( f\left( xw_{\left( 1 \right)} \right)w_{\left( 2 \right)} \right) \right)^{2}}
\end{equation}t

Here we have it, a way to quantify the cost of the neural network. This
function will be referred to as the \textbf{cost function}. Now, we have to solve
the problem of minimizing C. Will brute force work? It turns out,
no, because in a three-node neural network, there are millions of possible weights that we have to compute. But there are better solutions.

@@

Gradient descent~is "a~first-order~iterative~optimization~algorithm". To
find a~local/global minimum~of a function using gradient descent, we have to take
steps proportional to the gradient (or of the
approximate gradient) of the function at the current point.

The starting point of gradient descent in our case is random, since the initial set of weights are randomly generated. Then, the algorithm can take steps toward the minimum The process can be seen as a ball rolling down a
hill\footnote{\url{https://iamtrask.github.io/2015/07/27/python-network-part2/}
  by Andrew}
and trying to find the lowest point. Note that actual physics doesn't
apply here and we will define our own movement of the ball.

@@

\emph{Phase 1: Propagation: Each propagation
includes forward and backward propagation:}
\footnote{The steps are referenced from
Wikipedia who referenced:
\href{http://numericinsight.com/uploads/A_Gentle_Introduction_to_Backpropagation.pdf}{A
Gentle Introduction to Backpropagation - An intuitive tutorial by
Shashi Sathyanarayana}~The article contains pseudocode ("Training
Wheels for Training Neural Networks") for implementing the algorithm.}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Forward propagation of a training pattern's input through the neural
  network in order to generate the network's output value(s). 
\item
  Backward propagation of the propagation's output activations through
  the neural network using the target result in order to
  calculate the deltas (the difference between targeted and actual
  outputs) of all output and hidden neurons.
\end{enumerate}

\emph{
Phase 2: Weight update: For each weight:}

@@

Backpropagation is based around four fundamental equations. Together,
those equations give us a way of computing both the error
\(\delta^{L}\)and the gradient cost of the function. \footnote{The equations were
  created by Michael A. Neilson "Neural Networks and Deep Learning",
  Determination Press, 2015~} I didn't create these equations, in fact, a machine learning expert explained these commonly accepted equations in his book -- Neural Networks and Deep Learning.

@@

The first term on the right, \(\frac{\partial J}{\partial a_{J}^{L}}\)
measures how fast the cost is changing as a function of
j\textsuperscript{th} output activation. In simpler terms, this term calculates the derivative of C (cost) with respect to \(a\) (equation 2) at layer \(J\). The terms in (BP1) are easily
calculated. We computed \(z_{J}^{L}\) during forward feeding. Depending on the cost function, in our case, the quadratic
cost function is relatively easy to compute.

Equation (BP1) is a good expression, however, it's not matrix
based, form that backpropagation desires. But, we can transform it into a matrix form, which becomes: 
\begin{equation}
	\delta_{J}^{L} = (a_{(l)} - y)(\sigma'(z_{(l)})\
\end{equation}

@@

In the equation above, \(\left( w^{l + 1} \right)^{T}\) is the transpose of the weight
matrix \(\left( w^{l + 1} \right)\) for the
\emph{(l+1)\textsuperscript{th}} layer. We can calculate the error
\(\delta^{l + 1}\) at the \emph{(l+1)\textsuperscript{th }}layer by using (BP1). When
the transpose weight matrix is applied \(\left( w^{l + 1} \right)^{T}\),
we can think of this as moving the error backward through the network,
giving us some sort of measure of the error at the output of the
l\textsuperscript{th} layer. Finally, we take the Hadamard
product\(\ \bigodot\sigma'(z^{l})\).

Micheal Nilsson praises the power of (BP1) and (BP2): "By combining (BP1) with (BP2), the error \(\delta^{l}\) for any
layer in the network can by computed." We start by using~(BP1)~to
compute~\(\delta^{l}\), then apply Equation~(BP2)~to
compute~\(\delta^{L - 1}\), then Equation~(BP2)~again to
compute~\(\delta^{L - 2}\), and so on, all the way back through the
network. In the subsection below, there will be an example of using (BP1) with (BP2) in a simple three layer neural network.

@@

\subsubsection{An equation for the rate of change of the cost with respect to any weights in the network}
\begin{equation} \tag{BP4}
	\frac{\partial C}{\partial w^l_{jk}} = a^{l-1}_{k} \delta^l_{j}
\end{equation}

This equation can help us to compute the partial derivative \(\frac{\partial C}{\partial w^l_{jk}}\) in terms of the quantity \(\delta^l\) and \(a^{l-1}\). It can be rewritten in a less index-heavy notation:
\begin{equation}
	\frac{\partial C}{\partial w} = a_{in}\delta_{out}\
\end{equation}

Equation (6) is another critical component of backpropagation. Now, we know exactly how much to change the cost regarding the weights. If we apply the gradient to current weights, then, cost of the network should decrease, and we are closer to the global minimum of the cost function. 
\subsection{Backpropagation with simple neural network example}
